---
title: "section1"
author: "Helen Simpson"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

library(tidyverse)
library(ggplot2)
library(usmap)
library(lmtest)
library(broom)

popvote_df <- read_csv("./popvote_1948-2016.csv")
pvstate_df <- read_csv("./popvote_bystate_1948-2016.csv")

#ToDo: writeup
```

```{r extension.3, echo=FALSE}

#code from section, corrected so that the win margins map shows all maps since 1980 instead of just 2000
## map: win-margins
# pv_margins_map <- pvstate_df %>%
#     filter(year >= 1980) %>%
#     mutate(win_margin = (R_pv2p-D_pv2p))
# 
# plot_usmap(data = pv_margins_map, regions = "states", values = "win_margin") +
#   facet_wrap(facets = year ~.) +
#     scale_fill_gradient2(
#       high = "red", 
#       mid = scales::muted("purple"), ##TODO: purple or white better?
#       #mid = "white",
#       low = "blue", 
#       breaks = c(-50,-10,0,10,50), 
#       limits = c(-50,50),
#       name = "win margin"
#     ) +
#     theme_void()

#extension 3: change the formula from win margin to swing margin
#swing formula: D proportion of 2 party vote this year - D proportion of 2 party vote 4 years ago
#I also changed the formatting a little: I had to reset the gradient bc the difference ranges from about -17% to 17%. I also added axis labels
pv_change_map <- pvstate_df %>% 
  group_by(state) %>% arrange(year) %>% 
  mutate(prev_R = dplyr::lag(R), 
         prev_D = dplyr::lag(D),
         D_twoparty = 100* D / (R + D), #this should be the same as D_pv2p
         prev_D_twoparty = 100 * prev_D / (prev_D + prev_R),
         diff = D_twoparty - prev_D_twoparty) %>% 
  filter(year >= 1980) #if you do this before the mutate(), 1980 will have all N/As

extension3_plot <- plot_usmap(data = pv_change_map, regions = "states", values = "diff") +
  facet_wrap(facets = year ~.) +
    scale_fill_gradient2(high = "red", 
                         mid = scales::muted("purple"), ##TODO: purple or white better?
                         #mid = "white",
                         low = "blue", 
                         breaks = c(-20,-10,0,10,20), 
                         limits = c(-20,20),
                         name = "Change") +
  theme_void() +
  labs(title = "Change in percent of two-party vote from previous presidential election", 
       subtitle  = "Blue states voted more Democratic, red states voted more Republican")

ggsave("docs/images/extension3.png", plot = extension3_plot, height = 6, width = 14)


# ## map grid
# pv_map_grid <- pvstate_df %>%
#     filter(year >= 1980) %>%
#     mutate(winner = ifelse(R > D, "republican", "democrat"))
# 
# plot_usmap(data = pv_map_grid, regions = "states", values = "winner", color = "white") +
#     facet_wrap(facets = year ~.) + ## specify a grid by year
#     scale_fill_manual(values = c("blue", "red"), name = "PV winner") +
#     theme_void() +
#     theme(strip.text = element_text(size = 12),
#           aspect.ratio = 1)
# 
# ggsave("PV_states_historical.png", height = 3, width = 8)
```

```{r states.movement.charts, include=FALSE}
#copied from the code for section 1

pvstate_df$vote_margin <- pvstate_df$R_pv2p - pvstate_df$D_pv2p

pvstate_df %>% 
  ## subset data
  filter(state %in% c("California","Arizona", "Alaska")) %>%
  # filter(year >= 2000) %>%
  ## pipe into ggplot()
  ggplot(aes(x=year, y=vote_margin, color=vote_margin)) + 
  ## specify a grid by state
  facet_wrap(. ~ state) + 
  ## add plot elements
  geom_hline(yintercept=0,color="gray") +
  geom_line(size=2) + 
  geom_point(size=6) +
  ## specify scale colors
  scale_colour_gradient(low = "blue", high = "red") +
  scale_fill_gradient(low = "blue", high = "red") +
  ## specify titles, labels
  xlab("") +
  ylab("Republican vote-share margin") + 
  ggtitle("Swing states that moved sharply to the left in 2016") +
  ## switch position of x-axis and y-axis
  coord_flip() +
  ## make x-axis (year) run from top to bottom
  scale_x_reverse(breaks=unique(pvstate_df$year)) +
  theme_minimal() + 
  theme(panel.border    = element_blank(),
        plot.title      = element_text(size = 20, hjust = 0.5, face="bold"), 
        legend.position = "none",
        axis.title      = element_text(size=18),
        axis.text.x     = element_text(angle = 45, hjust = 1),
        axis.text       = element_text(size = 18),
        strip.text      = element_text(size = 18, face = "bold"))

```

-Inspired by this graphic from section
Let's start out with a basic question: how does partisanship affect states over time? Over this 68-year dataset, do states revert to the nationwide mean? Do states have individual means that they revert to?
-Why does reversion to the mean occur?
-Two potential processes: reversion to the mean and changing means over time. Which predicts better?

```{r}
#create a new variable for lagged R vote share. This is also done above under extension3, but that's inside someone else's code and it's cleaner do it again in my own style here.
pvstate_df <- pvstate_df %>% 
  group_by(state) %>% arrange(year) %>% 
  mutate(prev_Rvote = dplyr::lag(R_pv2p))

#make graph -- see example in The Gamble Ch. 7, p. 191
across_elections_plot <- pvstate_df %>% ggplot(aes(x = prev_Rvote, y = R_pv2p, color = year)) + 
  geom_point() + 
  scale_color_gradient("Year", low = "gray", high = "red") + 
  geom_abline() + 
  theme_classic() + 
  labs(x = "Republican vote in the previous election", y = "Republican two-party vote share in this election", title = "State-level Republican two-party vote share across elections", caption = "The black line is x = y.")

ggsave("docs/images/across_elections.png", plot = across_elections_plot, height = 4.5, width = 6.5)

```

```{r autocorrelation, echo=FALSE}
#run durbin-watson autocorrelation tests and plot results

#cleans n/a, selects relevant variables. not sure if this is still needed but it won't hurt. I added this stuff in response to various errors that earlier approaches were throwing.
pvstate_nona <- pvstate_df %>% 
  mutate(fake_var = 1) %>% #it won't let me run an empty regression - so I'll regress against a constant
  filter(!is.na(R_pv2p)) %>% 
  select(state, year, R_pv2p, fake_var)

#the regression against a constant idea didn't end up working bc the dwtest function didn't like it. I'll keep the variable here for posterity

#run dw tests on trendlines by state using dplyr::do
#do requires output to be dfs, so tidy (which turns test outputs into dfs). if you come across this problem again, see this page https://github.com/tidyverse/dplyr/issues/397
test_results = pvstate_nona %>% group_by(state) %>% 
  do(tidy(dwtest(lm(R_pv2p ~ year, data = .), alternative = c("two.sided"))))

#plot results of dw tests
autocorr_plot <- test_results %>% 
  ggplot(aes(x = statistic)) + 
  geom_histogram(bins = 10) + 
  theme_classic() +
  labs(x = "Durbin-Watson statistic", 
       y = "Frequency", 
       title = "Durbin-Watson Tests on State-level Results")

ggsave("docs/images/autocorr.png", plot = autocorr_plot, height = 2.5, width = 4)
     
```

```{r density, echo=FALSE}
#plot histograms of R vote by state for each election year in one diagram
density_by_year_plot <- pvstate_df %>% ggplot(aes(x = R_pv2p)) + facet_wrap(. ~year) + geom_density() + theme_classic() + labs(x = "Republican share of two-party vote by state", y = "Frequency", title = "How the distribution of Republican state-level vote share has changed over time")

ggsave("docs/images/density_by_year.png", plot = density_by_year_plot, height = 4, width = 7.25)

```