---
title: "section1"
author: "Helen Simpson"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

library(tidyverse)
library(ggplot2)
library(usmap)

popvote_df <- read_csv("./popvote_1948-2016.csv")
pvstate_df <- read_csv("./popvote_bystate_1948-2016.csv")

#ToDo: extension 3 map, autocorrelation test, changes in mean/ variance over time histograms 
```
```{r extension.3, echo=FALSE}

```

```{r}
#code from section, corrected so that the win margins map shows all maps since 1980 instead of just 2000
## map: win-margins
pv_margins_map <- pvstate_df %>%
    filter(year >= 1980) %>%
    mutate(win_margin = (R_pv2p-D_pv2p))

plot_usmap(data = pv_margins_map, regions = "states", values = "win_margin") +
  facet_wrap(facets = year ~.) +
    scale_fill_gradient2(
      high = "red", 
      mid = scales::muted("purple"), ##TODO: purple or white better?
      #mid = "white",
      low = "blue", 
      breaks = c(-50,-10,0,10,50), 
      limits = c(-50,50),
      name = "win margin"
    ) +
    theme_void()

#extension 3: change the formula from win margin to swing margin - these will not run yet bc I haven't figured out the formula yet! I'm taking a break now.
# pv_margins_map <- pvstate_df %>%
#     filter(year >= 1972) %>%
#     mutate(swing_margin = ())

plot_usmap(data = pv_margins_map, regions = "states", values = "swing_margin") +
  facet_wrap(facets = year ~.) +
    scale_fill_gradient2(
      high = "red", 
      mid = scales::muted("purple"), ##TODO: purple or white better?
      #mid = "white",
      low = "blue", 
      breaks = c(-50,-10,0,10,50), 
      limits = c(-50,50),
      name = "Swing margin"
    ) +
    theme_void()


# ## map grid
# pv_map_grid <- pvstate_df %>%
#     filter(year >= 1980) %>%
#     mutate(winner = ifelse(R > D, "republican", "democrat"))
# 
# plot_usmap(data = pv_map_grid, regions = "states", values = "winner", color = "white") +
#     facet_wrap(facets = year ~.) + ## specify a grid by year
#     scale_fill_manual(values = c("blue", "red"), name = "PV winner") +
#     theme_void() +
#     theme(strip.text = element_text(size = 12),
#           aspect.ratio = 1)
# 
# ggsave("PV_states_historical.png", height = 3, width = 8)
```

```{r states.movement.charts, include=FALSE}
#copied from the code for section 1

pvstate_df$vote_margin <- pvstate_df$R_pv2p - pvstate_df$D_pv2p

pvstate_df %>% 
  ## subset data
  filter(state %in% c("New Hampshire","Illinois", "Massachusetts")) %>%
  # filter(year >= 2000) %>%
  ## pipe into ggplot()
  ggplot(aes(x=year, y=vote_margin, color=vote_margin)) + 
  ## specify a grid by state
  facet_wrap(. ~ state) + 
  ## add plot elements
  geom_hline(yintercept=0,color="gray") +
  geom_line(size=2) + 
  geom_point(size=6) +
  ## specify scale colors
  scale_colour_gradient(low = "blue", high = "red") +
  scale_fill_gradient(low = "blue", high = "red") +
  ## specify titles, labels
  xlab("") +
  ylab("Republican vote-share margin") + 
  ggtitle("Swing states that moved sharply to the left in 2016") +
  ## switch position of x-axis and y-axis
  coord_flip() +
  ## make x-axis (year) run from top to bottom
  scale_x_reverse(breaks=unique(pvstate_df$year)) +
  theme_minimal() + 
  theme(panel.border    = element_blank(),
        plot.title      = element_text(size = 20, hjust = 0.5, face="bold"), 
        legend.position = "none",
        axis.title      = element_text(size=18),
        axis.text.x     = element_text(angle = 45, hjust = 1),
        axis.text       = element_text(size = 18),
        strip.text      = element_text(size = 18, face = "bold"))

```
-Inspired by this graphic from section
Let's start out with a basic question: how does partisanship affect states over time? Over this 68-year dataset, do states revert to the nationwide mean? Do states have individual means that they revert to?
-Why does reversion to the mean occur?
-Two potential processes: reversion to the mean and changing means over time. Which predicts better?
```{r}
pvstate_df <- pvstate_df %>% rowwise() %>% mutate(previous_year = toString(year - 4)) %>% group_by(state, year)

pvstate_df <- pvstate_df %>% 
  group_by(state) %>% arrange(year) %>% 
  mutate(prev_Rvote = dplyr::lag(R_pv2p))

pvstate_df %>% ggplot(aes(x = prev_Rvote, y = R_pv2p, color = year)) + 
  geom_point() + 
  scale_color_gradient("Year", low = "gray", high = "red") + 
  geom_abline() + 
  theme_classic() + 
  labs(x = "Republican vote in the previous election", y = "Republican two-party vote share in this election", title = "State-level Republican two-party vote share across elections", caption = "The black line is x = y.")

```
